---
title: "Election2016"
author: "MukilaR"
date: "10 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The data used in this project is a public polling data organized by FiveThirtyEight for the 2016 presidential election. The table includes results for national polls, as well as state polls, taken in the year before the election. It contains 12625 rows and 27 columns.

```{r loadingdata, echo=FALSE}
library(dplyr)
library(dslabs)
data("polls_us_election_2016")
```

Here we will use all the national polls that ended within a few weeks before the election.

Assuming there are only two candidates, for each poll in the polling data set, use the CLT to create a 95% confidence interval for the spread. 
```{r}

# Creating a table called `polls` that filters by  state, date, and reports the spread
polls <- polls_us_election_2016 %>% 
  filter(state != "U.S." & enddate >= "2016-10-31") %>% 
  mutate(spread = rawpoll_clinton/100 - rawpoll_trump/100)

# An object called `cis` contains the columns for the lower and upper confidence intervals. 
cis <- polls %>% mutate(X_hat = (spread+1)/2, se = 2*sqrt(X_hat*(1-X_hat)/samplesize), 
                 lower = spread - qnorm(0.975)*se, upper = spread + qnorm(0.975)*se) %>%
  select(state, startdate, enddate, pollster, grade, spread, lower, upper)
cis

```
Now we add actual results calculated to the previous table and determine how often the 95% confidence interval includes the actual result.
```{r}
# Adding the actual results to the `cis` data set
add <- results_us_election_2016 %>% mutate(actual_spread = clinton/100 - trump/100) %>% select(state, actual_spread)
cis <- cis %>% mutate(state = as.character(state)) %>% left_join(add, by = "state")

# `p_hits` summarizes the proportion of confidence intervals that contain the actual value. 
p_hits<-cis %>% mutate(hit = lower <= actual_spread & upper >= actual_spread) %>%
summarize(proportion_hits=mean(hit))
p_hits
```
Then we find the proportion of hits for each pollster showing only pollsters with more than 5 polls and ordering them from best to worst. Also we show the number of polls conducted by each pollster and the FiveThirtyEight grade of each pollster.
```{r}
p_hits <- cis %>% mutate(hit = lower <= actual_spread & upper >= actual_spread) %>% 
  group_by(pollster) %>%
  filter(n() >=  5) %>%
  summarize(proportion_hits = mean(hit), n = n(), grade = grade[1]) %>%
  arrange(desc(proportion_hits))
p_hits
```
Repeating the previous code, but instead of pollster, we stratify by state. Here we cannot include grades.
```{r}
p_hits <- cis %>% mutate(hit = lower <= actual_spread & upper >= actual_spread) %>% 
  group_by(state) %>%
  filter(n() >=  5) %>%
  summarize(proportion_hits = mean(hit), n = n()) %>%
  arrange(desc(proportion_hits)) 
p_hits
```
A barplot is created for this new state table.
```{r fig.height = 9, fig.width = 7, echo=FALSE}
p_hits %>% mutate(state = reorder(state, proportion_hits)) %>%
  ggplot(aes(state, proportion_hits)) + 
  geom_bar(width = 0.5, stat = "identity", fill="darkblue") +
  coord_flip()
```

